---
title: "Optimizing Direct Marketing Campaign Success in Financial Services"
subtitle: "A Data Analytics Approach"
author: "Ravina Gaikwad"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
    highlight: tango
    number_sections: true
---

```{r setup, include=FALSE}
# ============================================================
# GLOBAL SETTINGS
# This chunk runs first and sets options for ALL chunks below.
# include=FALSE means this chunk won't appear in your output.
# ============================================================

knitr::opts_chunk$set(
  echo    = TRUE,    # Show your R code in the output
  warning = FALSE,   # Hide warnings (cleaner output)
  message = FALSE,   # Hide package loading messages
  fig.width  = 8,
  fig.height = 5
)
```

---

# Title

**Project Title:** Optimizing Direct Marketing Campaign Success in Financial Services

**Research Question:** Which customer characteristics and campaign factors best predict whether a client will subscribe to a bank term deposit?

**Student Name:** Your Name

**Course:** Your Course Name

**Date:** `r Sys.Date()`

---

# Problem Statement & Research Question

## The Problem

Financial institutions conducting direct marketing campaigns face the fundamental challenge of inefficient resource allocation and suboptimal conversion rates. Despite significant investment in outreach, most customers contacted do not subscribe to the product being offered.

## Research Question

> Is there a statistically significant relationship between customer demographic/behavioral characteristics and the likelihood of subscribing to a bank term deposit?

## Hypotheses

- **H₀ (Null):** Customer demographic characteristics, financial profile, and campaign factors have no significant effect on the likelihood of subscribing to a term deposit.
- **H₁ (Alternative):** At least one customer demographic, financial, or campaign variable significantly predicts the likelihood of subscribing to a term deposit.

---

## Relevance

This analysis is particularly valuable for small to medium-sized financial services businesses seeking data-driven growth strategies in competitive markets, where accurate customer targeting can significantly improve campaign ROI and reduce contact fatigue.

---

# Data Description & Acquisition

## Data Source

```{r load-packages}
# ============================================================
# LOAD PACKAGES
# Install any missing packages first by running in the Console:
#   install.packages("tidyverse")
#   install.packages("janitor")
#   install.packages("skimr")
#   install.packages("visdat")
#   install.packages("ggplot2")
#   install.packages("corrplot")
#   install.packages("car")
# ============================================================

library(tidyverse)   # Data manipulation and visualization (dplyr, ggplot2, etc.)
library(janitor)     # Cleans messy column names
library(skimr)       # Quick data summaries
library(visdat)      # Visualize missing data
library(corrplot)    # Correlation plots
library(car)         # Assumption testing (VIF, Levene's)
```

```{r import-data}
# ============================================================
# IMPORT DATA
# This dataset uses semicolons (;) as separators — not commas!
# ============================================================

# bank <- read_csv(
 # "/Users/rgaikw468/Downloads/bank+marketing/bank-additional/bank-additional.csv",
  #col_types = cols(.default = "c"),   # Read everything as character first (safer)
  #locale = locale(decimal_mark = ".")  
# ) |>
  # Wait — this file uses semicolons, so we use read_delim instead:
  # Re-run with the correct function below if you get errors.
  #clean_names()                        # Standardizes column names to snake_case

bank <- read_delim("/Users/rgaikw468/Downloads/bank+marketing/bank-additional/bank-additional.csv", delim = ";") |> clean_names()

# Preview the first few rows
glimpse(bank)
```

## Sample Size

```{r sample-size}
# ============================================================
# SAMPLE SIZE SUMMARY
# "Final N" will be used from what we get after data screening
# ============================================================

cat("Initial N (rows):", nrow(bank), "\n")
cat("Number of variables:", ncol(bank), "\n")
```

- **Data Source:** UCI Machine Learning Repository — Portuguese Banking Institution
  (Moro, Rita, & Cortez, 2014)
- **Initial N:** ~4,119 observations (10% stratified sample of full dataset)
- **Final N:** 3,829 (after removing 290 multivariate outliers via Mahalanobis distance)
- **Time Period:** May 2008 – November 2010

## Variables

- **Predictors (IVs):** Age, job, marital status, education, credit default, account balance, housing loan, personal loan, contact type, day of week, month, campaign contacts, previous contacts, previous outcome, and 5 economic indicators
- **Outcome (DV):** `y` — Has the client subscribed to a term deposit? (yes / no)

```{r variable-overview}
# Quick summary of all variables
skim(bank)
```

---

# Data Screening

> **The "Pre-Flight Checklist"** — Before running any analysis, we must verify our data is clean, complete, and meets statistical assumptions.

## Step 1: Fix Column Types

```{r fix-types}
# ============================================================
# CONVERT COLUMNS TO CORRECT TYPES
# Right now everything is character — we need to convert 
# numeric columns to numeric and factors to factors.
# ============================================================

bank_clean <- bank |>
  mutate(
    # --- Numeric variables ---
    age              = as.numeric(age),
    duration         = as.numeric(duration),
    campaign         = as.numeric(campaign),
    pdays            = as.numeric(pdays),
    previous         = as.numeric(previous),
    emp_var_rate     = as.numeric(emp_var_rate),
    cons_price_idx   = as.numeric(cons_price_idx),
    cons_conf_idx    = as.numeric(cons_conf_idx),
    euribor3m        = as.numeric(euribor3m),
    nr_employed      = as.numeric(nr_employed),
    
    # --- Outcome variable: convert to binary (1 = yes, 0 = no) ---
    y = if_else(y == "yes", 1L, 0L),
    
    # --- Categorical / Factor variables ---
    job       = as.factor(job),
    marital   = as.factor(marital),
    education = as.factor(education),
    default   = as.factor(default),
    housing   = as.factor(housing),
    loan      = as.factor(loan),
    contact   = as.factor(contact),
    month     = as.factor(month),
    day_of_week = as.factor(day_of_week),
    poutcome  = as.factor(poutcome)
  )

glimpse(bank_clean)
```

## Step 2: Accuracy Check

```{r accuracy-check}
# ============================================================
# CHECK FOR IMPOSSIBLE OR SUSPICIOUS VALUES
# Looking for negative ages, impossible durations, etc.
# ============================================================

# Summary of numeric variables
bank_clean |>
  select(where(is.numeric)) |>
  summary()
```

```{r accuracy-notes}
# ============================================================
# NOTE ON 'UNKNOWN' CATEGORIES
# Several categorical variables contain "unknown" as a level.
# These are effectively a form of missing data — we'll address
# them in the missing data section below.
# ============================================================

# Count "unknown" values in each categorical column
bank_clean |>
  select(where(is.factor)) |>
  summarise(across(everything(), ~ sum(. == "unknown", na.rm = TRUE))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "n_unknown") |>
  filter(n_unknown > 0)
```

**Findings from the accuracy check:**

All numeric variables fell within plausible ranges. Notably:

- `age` ranges from 18 to 88 — both values are realistic and were retained.
- `duration` has a minimum of 0 seconds, which is technically possible (contact made but immediately ended) and was retained.
- `campaign` has a maximum of 35 contacts — extremely high but not impossible; flagged as a potential outlier in univariate screening below.
- `pdays` is coded as 999 for the vast majority of clients, meaning they were not previously contacted. This is a design choice in the dataset, not an error — it was retained as-is since it encodes meaningful information.

**"Unknown" categories detected in 6 variables** — these act as implicit missing data and were treated as a separate factor level rather than true NAs, which is appropriate for logistic regression:

| Variable  | N Unknown |
|-----------|-----------|
| job       | 39        |
| marital   | 11        |
| education | 167       |
| default   | 803       |
| housing   | 105       |
| loan      | 105       |

The `default` variable has the highest proportion of unknowns (803 out of 4,119, ~19.5%). This was retained as a factor level since true credit default information is often legally restricted from direct use in marketing models.

## Step 3: Missing Data

```{r missing-data}
# ============================================================
# CHECK FOR MISSING VALUES (NAs)
# using visdat which will give us a visual map; using colSums will give us exact counts.
# ============================================================

# Visual map of missingness
vis_miss(bank_clean)

# Count of missing values per column
colSums(is.na(bank_clean))
```

```{r missing-mechanism}
# ============================================================
# MISSING DATA STRATEGY
# If < 5% missing: Listwise deletion is acceptable.
# If >= 5% missing: Consider multiple imputation (mice package).
# 
# Calculate % missing per variable:
# ============================================================

missing_pct <- colSums(is.na(bank_clean)) / nrow(bank_clean) * 100
missing_pct[missing_pct > 0]
```

- **Mechanism Identified:** No true NAs were found in any column (`colSums(is.na())` returned 0 for all 21 variables). The dataset is structurally complete. The only form of "missingness" present is encoded as the string `"unknown"` in six categorical variables, which was retained as a valid factor level rather than imputed.
- **Strategy:** No imputation or listwise deletion was required. `drop_na()` confirmed 0 rows were removed (N remained at 4,119 after this step).

```{r handle-missing}
# ============================================================
# HANDLE MISSING VALUES
# If listwise deletion is appropriate:
# ============================================================

bank_screened <- bank_clean |>
  drop_na()    # Removes rows with ANY missing value

cat("Rows after dropping NAs:", nrow(bank_screened), "\n")
```

## Step 4: Outlier Detection

```{r outliers-univariate}
# ============================================================
# UNIVARIATE OUTLIERS — Z-Score Method
# Flag observations where Z > 3.29 (p < .001, two-tailed)
# ============================================================

numeric_vars <- bank_screened |> select(where(is.numeric), -y)

z_scores <- numeric_vars |>
  mutate(across(everything(), ~ (. - mean(., na.rm = TRUE)) / sd(., na.rm = TRUE)))

# Count outliers per variable
outlier_counts <- z_scores |>
  summarise(across(everything(), ~ sum(abs(.) > 3.29, na.rm = TRUE)))

print(outlier_counts)
```

```{r outliers-multivariate}
# ============================================================
# MULTIVARIATE OUTLIERS — Mahalanobis Distance
# Flag observations where p < .001
# ============================================================

numeric_matrix <- numeric_vars |> drop_na() |> as.matrix()

mahal_dist <- mahalanobis(
  numeric_matrix,
  colMeans(numeric_matrix, na.rm = TRUE),
  cov(numeric_matrix, use = "complete.obs")
)

# Calculate p-values
mahal_p <- pchisq(mahal_dist, df = ncol(numeric_matrix), lower.tail = FALSE)

# How many are outliers?
n_mahal_outliers <- sum(mahal_p < .001)
cat("Multivariate outliers (Mahalanobis p < .001):", n_mahal_outliers, "\n")

# Remove multivariate outliers
bank_final <- bank_screened[mahal_p >= .001, ]
cat("Final N after outlier removal:", nrow(bank_final), "\n")
```

- **Univariate Outliers (Z > 3.29):** Detected across several variables — 24 in `age`, 74 in `duration`, 87 in `campaign`, 160 in `pdays`, and 121 in `previous`. Economic indicator variables (`emp_var_rate`, `cons_price_idx`, `cons_conf_idx`, `euribor3m`, `nr_employed`) had zero univariate outliers, reflecting their relatively stable distributions across the dataset.
- **Multivariate Outliers (Mahalanobis p < .001):** 290 observations were flagged as multivariate outliers when examining all numeric predictors simultaneously.
- **Decision:** The 290 multivariate outliers were **removed**, as Mahalanobis distance is the gold-standard method for detecting cases that are unusual across the combined pattern of all predictors — even if they appear normal on any single variable in isolation. Final N after removal: **3,829**.

---

# Exploratory Data Analysis (EDA)

> All visualizations follow Tufte's principles: maximized data-ink ratio, no chartjunk.

## Distribution of Outcome Variable

```{r eda-outcome}
# ============================================================
# SUBSCRIPTION RATE
# How many clients subscribed (y = 1) vs. did not (y = 0)?
# ============================================================

bank_final |>
  mutate(Subscribed = if_else(y == 1, "Yes", "No")) |>
  count(Subscribed) |>
  mutate(Percent = round(n / sum(n) * 100, 1)) |>
  ggplot(aes(x = Subscribed, y = n, fill = Subscribed)) +
  geom_col(width = 0.5) +
  geom_text(aes(label = paste0(Percent, "%")), vjust = -0.5, size = 4) +
  scale_fill_manual(values = c("Yes" = "#2E86AB", "No" = "#A8DADC")) +
  labs(
    title = "Term Deposit Subscription Rate",
    x = "Subscribed",
    y = "Count",
    caption = "Source: UCI Bank Marketing Dataset (Moro et al., 2014)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

## Age Distribution by Subscription

```{r eda-age}
# ============================================================
# AGE DISTRIBUTION
# Do subscribers tend to be older or younger?
# ============================================================

bank_final |>
  mutate(Subscribed = if_else(y == 1, "Yes", "No")) |>
  ggplot(aes(x = age, fill = Subscribed)) +
  geom_histogram(binwidth = 5, alpha = 0.7, position = "identity") +
  scale_fill_manual(values = c("Yes" = "#2E86AB", "No" = "#E63946")) +
  labs(
    title = "Age Distribution by Subscription Status",
    x = "Age",
    y = "Count",
    fill = "Subscribed"
  ) +
  theme_minimal()
```

## Subscription Rate by Job Type

```{r eda-job}
# ============================================================
# SUBSCRIPTION RATE BY JOB
# Finding out job categories that have the highest conversion rates.
# ============================================================

bank_final |>
  mutate(Subscribed = if_else(y == 1, "Yes", "No")) |>
  group_by(job, Subscribed) |>
  summarise(n = n(), .groups = "drop") |>
  group_by(job) |>
  mutate(pct = n / sum(n) * 100) |>
  filter(Subscribed == "Yes") |>
  ggplot(aes(x = reorder(job, pct), y = pct)) +
  geom_col(fill = "#2E86AB") +
  coord_flip() +
  labs(
    title = "Subscription Rate by Job Type",
    x = "Job",
    y = "Subscription Rate (%)"
  ) +
  theme_minimal()
```

## Call Duration vs. Subscription

```{r eda-duration}
# ============================================================
# CALL DURATION BOXPLOT
# Longer calls may indicate higher engagement / interest.
# Note: Duration is known AFTER the call, so it cannot be 
# used as a predictor in real-time models. We include it 
# here for EDA only.
# ============================================================

bank_final |>
  mutate(Subscribed = if_else(y == 1, "Yes", "No")) |>
  ggplot(aes(x = Subscribed, y = duration, fill = Subscribed)) +
  geom_boxplot(alpha = 0.7, outlier.alpha = 0.3) +
  scale_fill_manual(values = c("Yes" = "#2E86AB", "No" = "#E63946")) +
  labs(
    title = "Call Duration by Subscription Status",
    x = "Subscribed",
    y = "Duration (seconds)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

**EDA Observations:**

- **Subscription rate is highly imbalanced:** Only ~11% of clients in the cleaned dataset (N = 3,829) subscribed to the term deposit. This class imbalance is important to keep in mind when interpreting the logistic regression — the model must work harder to detect the minority class (subscribers).
- **Age distribution is right-skewed:** Most clients are concentrated between ages 30–50, with a long tail toward older ages (up to 88). Both subscribers and non-subscribers share a similar age distribution, though subscribers show a slightly higher density in the 25–35 and 60+ age ranges, suggesting students and retirees may be more receptive.
- **Job type strongly differentiates subscription rates:** Students and retired individuals show the highest subscription rates among all job categories, while blue-collar workers and services staff show the lowest. This aligns with the economic intuition that individuals with more financial flexibility or time are more likely to invest in term deposits.
- **Call duration is a strong EDA signal (but not a usable predictor):** Subscribers had substantially longer average call durations than non-subscribers. However, duration is only known after the call ends and therefore cannot be used as a predictor in a real-time targeting model — it is included here for descriptive insight only and was excluded from the final regression.

---

# Methods & Techniques

## Software

- **Language:** R (version `r R.version$major`.`r R.version$minor`)
- **IDE:** RStudio
- **Key Packages:** `tidyverse`, `janitor`, `skimr`, `car`, `visdat`

## Assumption Testing

```{r assumption-multicollinearity}
# ============================================================
# ASSUMPTION: NO MULTICOLLINEARITY
# We'll run a preliminary logistic regression and check VIF.
# VIF > 10 = serious multicollinearity problem.
# ============================================================

# Preliminary model for assumption checking
# (We'll build the proper model in Phase 4)
prelim_model <- glm(
  y ~ age + duration + campaign + previous + euribor3m + nr_employed,
  data   = bank_final,
  family = binomial(link = "logit")
)
```

```{r assumption-vif-correlation}
# Check Variance Inflation Factor (VIF)
vif(prelim_model)
# ============================================================
# ASSUMPTION: ADDITIVITY (MULTICOLLINEARITY)
# Two approaches:
#   A) Correlation matrix — flag pairs where |r| >= 0.80
#   B) VIF — flag variables where VIF >= 10
#      Ideal: average VIF < 2
# ============================================================

# --- 2A: Correlation Matrix (numeric predictors only) ---

numeric_preds <- bank_final |>
  select(age, duration, campaign, previous, euribor3m, nr_employed,
         emp_var_rate, cons_price_idx, cons_conf_idx, pdays)

cor_matrix <- cor(numeric_preds, use = "complete.obs")

# Print rounded correlation matrix
cat("===== CORRELATION MATRIX =====\n")
print(round(cor_matrix, 2))

# Visualise it — easier to spot problem pairs at a glance
corrplot(
  cor_matrix,
  method  = "color",
  type    = "upper",
  addCoef.col = "black",
  number.cex  = 0.65,
  tl.cex      = 0.8,
  col    = colorRampPalette(c("#E63946", "white", "#2E86AB"))(200),
  title  = "Correlation Matrix — Numeric Predictors",
  mar    = c(0, 0, 1.5, 0)
)

# Flag any pairs that exceed r = 0.80
cat("\n===== HIGH CORRELATION PAIRS (|r| >= 0.80) =====\n")
high_cor <- which(abs(cor_matrix) >= 0.80 & upper.tri(cor_matrix), arr.ind = TRUE)
if (nrow(high_cor) == 0) {
  cat("None found — all pairs are below r = 0.80 ✅\n")
} else {
  for (i in 1:nrow(high_cor)) {
    r1 <- rownames(cor_matrix)[high_cor[i, 1]]
    r2 <- colnames(cor_matrix)[high_cor[i, 2]]
    cat(sprintf("  %s & %s: r = %.2f ⚠️\n", r1, r2, cor_matrix[high_cor[i,1], high_cor[i,2]]))
  }
}

# --- 2B: Variance Inflation Factor (VIF) ---
# Uses the prelim_model defined in Assumption 1 above.
# VIF >= 10 = serious problem
# Average VIF ideally < 2

vif_vals <- vif(prelim_model)

cat("===== VIF VALUES =====\n")
print(round(vif_vals, 3))

cat("\nAverage VIF:", round(mean(vif_vals), 3), "\n")

# Flag any violations
cat("\n===== VIF FLAGS =====\n")
if (any(vif_vals >= 10)) {
  cat("⚠️  Serious multicollinearity detected (VIF >= 10):", 
      names(vif_vals[vif_vals >= 10]), "\n")
} else {
  cat("No VIF >= 10 ✅\n")
}

if (mean(vif_vals) >= 2) {
  cat("⚠️  Average VIF >= 2 — moderate multicollinearity present. Interpret affected coefficients with caution.\n")
} else {
  cat("Average VIF < 2 ✅\n")
}
```
The correlation matrix reveals strong collinearity between the economic indicator variables — notably euribor3m and nr_employed (r ≈ −0.95). VIF values for these two variables are 8.34 and 8.64 respectively, which are below the critical threshold of 10 but elevated enough to warrant caution when interpreting their individual coefficients. No VIF exceeds 10, so no predictors were dropped on multicollinearity grounds alone.

```{r assumption-independence}
# ============================================================
# ASSUMPTION: INDEPENDENCE OF OBSERVATIONS
# Durbin-Watson test: values near 2 = no autocorrelation.
# Values < 1 or > 3 may be problematic.
# ============================================================

durbinWatsonTest(prelim_model)
```
The Durbin-Watson statistic of 1.876 is acceptably close to 2.0, indicating no meaningful autocorrelation in the residuals. The assumption of independence is satisfied.

## Primary Analysis: Logistic Regression

Logistic regression is the appropriate technique here because our outcome variable (`y`) is binary (subscribed: yes/no). The model estimates the **log-odds** of subscription as a function of predictor variables.

$$\log\left(\frac{p}{1-p}\right) = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \cdots + \beta_k X_k$$

---

# Challenges Encountered

## Data Hygiene

The dataset arrived in a largely tidy format — each row represented one client contact and each column represented one variable — so no reshaping with `pivot_longer()` was required. However, two hygiene challenges were encountered:

- **Semicolon delimiter:** The file uses `;` as a separator instead of the standard `,`. The initial `read_csv()` call failed to parse the file correctly and had to be replaced with `read_delim(..., delim = ";")`. This is a common beginner pitfall with European-format CSV files.
- **Column name standardization:** `clean_names()` from the `janitor` package was applied to convert all column names to `snake_case` (e.g., `emp.var.rate` became `emp_var_rate`). This prevents errors caused by dots in variable names when used inside `dplyr` pipelines.
- **All-character import:** Columns were initially read as character type using `col_types = cols(.default = "c")` to avoid R silently misclassifying variables. Manual type conversion was then applied deliberately in `bank_clean`.

## Screening Issues

- **"Unknown" as implicit missing data:** Six categorical variables — `job` (n=39), `marital` (n=11), `education` (n=167), `default` (n=803), `housing` (n=105), and `loan` (n=105) — contained `"unknown"` as a category. These were retained as valid factor levels rather than being imputed or dropped, because the fact that a value is unknown may itself carry predictive signal (particularly for `default`).
- **No structural NAs found:** `colSums(is.na())` returned 0 across all 21 columns, confirming no true missing values existed after type conversion.
- **Multicollinearity between economic indicators:** The preliminary VIF check flagged `euribor3m` (VIF = 8.34) and `nr_employed` (VIF = 8.64) as moderately collinear. These values are below the critical threshold of 10, so both were retained in the final model, but their high correlation with each other means their individual coefficients should be interpreted with caution.
- **Durbin-Watson result:** The Durbin-Watson statistic was 1.876, which is acceptably close to 2.0, indicating no meaningful autocorrelation in the residuals. The associated p-value of 0 reflects the large sample size making even small deviations statistically detectable — this is not a practical concern.

## Technical Roadblocks

- **Variable name mismatch:** The original template used `emp.var.rate`, `cons.price.idx`, etc. with dots, but `clean_names()` converted these to underscores (`emp_var_rate`, `cons_price_idx`). This caused errors in the `fix-types` chunk that referenced the old dot-notation names, which had to be updated manually.
- **Singularity warning in final model:** R flagged one coefficient as `NA` (`loanunknown`) due to perfect or near-perfect collinearity with another variable. This is because `loan` and `housing` both have identical `"unknown"` counts (105 each), suggesting the same rows carry that value — R automatically dropped the redundant coefficient. This is handled gracefully by `glm()` and does not invalidate the model.

---
# Statistical Analysis & Results

## Final Logistic Regression Model

```{r final-model}
# ============================================================
# FINAL LOGISTIC REGRESSION MODEL
# 
# TODO: Adjust the predictors below based on your EDA findings
# and theoretical reasoning. Remove variables with VIF > 10.
#
# NOTE: We exclude 'duration' from the final model because it
# is only known after the call ends — it cannot be used to
# predict subscription BEFORE contacting a customer.
# ============================================================

final_model <- glm(
  y ~ age + job + marital + education + default + housing + 
      loan + contact + month + campaign + poutcome + 
      euribor3m + nr_employed,
  data   = bank_final,
  family = binomial(link = "logit")
)

summary(final_model)
```

## Effect Sizes

```{r effect-sizes}
# ============================================================
# EFFECT SIZES FOR LOGISTIC REGRESSION
#
# 1. Odds Ratios (OR): exp(coef) — how much each predictor
#    multiplies the odds of subscribing.
#    OR > 1 = increases odds, OR < 1 = decreases odds.
#
# 2. Nagelkerke R²: overall model fit (0–1 scale).
#    Like R² in linear regression but for logistic models.
# ============================================================

# Odds Ratios with 95% Confidence Intervals
odds_ratios <- exp(cbind(OR = coef(final_model), confint(final_model)))
round(odds_ratios, 3)
```

```{r nagelkerke}
# ============================================================
# NAGELKERKE R² — Pseudo R-squared for logistic regression
# ============================================================

# Null model (intercept only)
null_model <- glm(y ~ 1, data = bank_final, family = binomial(link = "logit"))

# Nagelkerke R²
n <- nrow(bank_final)
nagelkerke_r2 <- (1 - exp((logLik(null_model) - logLik(final_model)) * (2/n))) /
                 (1 - exp(logLik(null_model) * (2/n)))

cat("Nagelkerke R²:", round(nagelkerke_r2, 3), "\n")
```
Nagelkerke R² = 0.179, indicating that the model explains approximately 17.9% of the variance in subscription outcomes. This is a modest but meaningful fit for a behavioural marketing dataset with many real-world confounders.

```{r chi-square-lrt}
# ============================================================
# OVERALL MODEL SIGNIFICANCE — Chi-Square Likelihood Ratio Test
# This is the logistic regression equivalent of the F-statistic
# in linear regression. It tests whether the full model with
# all predictors fits significantly better than a null model
# (intercept only — i.e., no predictors at all).
#
# How it works:
#   χ² = 2 × (log-likelihood of full model
#              − log-likelihood of null model)
#   df  = number of predictors added (full model df − null df)
#   p   < .05 means the model as a whole is significant.
# ============================================================

# --- Compute the likelihood ratio chi-square test ---
chi_sq_val <- 2 * (logLik(final_model) - logLik(null_model))
chi_sq_df  <- final_model$df.null - final_model$df.residual
chi_sq_p   <- pchisq(chi_sq_val, df = chi_sq_df, lower.tail = FALSE)

cat("===== OVERALL MODEL FIT — Likelihood Ratio Test =====\n\n")
cat("Null deviance     :", round(null_model$deviance, 2),
    "(df =", null_model$df.residual, ")\n")
cat("Residual deviance :", round(final_model$deviance, 2),
    "(df =", final_model$df.residual, ")\n")
cat("Chi-square (χ²)   :", round(chi_sq_val, 2), "\n")
cat("Degrees of freedom:", chi_sq_df, "\n")
cat("p-value           :", format.pval(chi_sq_p, digits = 3), "\n")
cat("Nagelkerke R²     :", round(nagelkerke_r2, 3), "\n\n")

# --- Plain-English interpretation ---
if (chi_sq_p < .001) {
  cat("Interpretation: The model is statistically significant (p < .001).\n")
  cat("Adding the predictors explains significantly more variance in\n")
  cat("subscription likelihood than the intercept-only baseline model.\n")
} else if (chi_sq_p < .05) {
  cat("Interpretation: The model is statistically significant (p < .05).\n")
} else {
  cat("Interpretation: The model is NOT significant (p > .05).\n")
  cat("The predictors do not improve fit over the null model.\n")
}
```
## The "Money Plot"

```{r money-plot}
# ============================================================
# MONEY PLOT: Predicted Probability of Subscription
# by Number of Employees (nr_employed) — the strongest
# continuous predictor in the final model (p < .001).
# Lower employment levels correspond to periods of looser
# monetary policy, which appear to drive higher subscription rates.
# ============================================================

# Create a prediction data frame
pred_data <- bank_final |>
  mutate(predicted_prob = predict(final_model, type = "response"))

# Plot predicted probabilities by a key continuous predictor
pred_data |>
  ggplot(aes(x = nr_employed, y = predicted_prob)) +
  geom_point(aes(color = as.factor(y)), alpha = 0.3, size = 1) +
  geom_smooth(method = "loess", se = TRUE, color = "#2E86AB") +
  scale_color_manual(
    values = c("0" = "#E63946", "1" = "#2E86AB"),
    labels = c("Did Not Subscribe", "Subscribed"),
    name   = "Actual Outcome"
  ) +
  labs(
    title    = "Predicted Probability of Subscription by Number of Employee",
    subtitle = "Logistic Regression Model",
    x        = "Number of Employee",
    y        = "Predicted Probability of Subscription",
    caption  = "Source: UCI Bank Marketing Dataset (Moro et al., 2014)"
  ) +
  theme_minimal()
```

## APA-Style Results Write-Up

A binary logistic regression was conducted to predict the likelihood of term deposit subscription from customer demographic, financial, and campaign characteristics (N = 3,829). The overall model was statistically significant, χ²(42, N = 3,829) = 309.6, p < .001, Nagelkerke R² = .179, indicating that the predictors collectively explained approximately 17.9% of the variance in subscription outcomes.

Several predictors emerged as statistically significant. The number of employees in the economy (`nr_employed`) was the strongest continuous predictor (OR = 0.987, 95% CI [0.980, 0.993], p < .001), such that periods with fewer employed workers — reflecting looser economic conditions — were associated with higher odds of subscription. Contact via telephone compared to cellular was associated with significantly lower odds of subscribing (OR = 0.547, 95% CI [0.348, 0.836], p = .007), suggesting that the channel of outreach meaningfully affects campaign success. Campaign timing also mattered: contacts made in March (OR = 3.604, 95% CI [1.566, 8.299], p = .002), June (OR = 2.140, 95% CI [1.248, 3.723], p = .006), and December (OR = 4.001, 95% CI [1.278, 12.996], p = .018) were associated with substantially higher subscription odds relative to the reference month. Additionally, clients with a non-existent prior campaign outcome (`poutcomenonexistent`) showed higher odds of subscribing compared to those with a failure outcome (OR = 1.532, 95% CI [1.041, 2.294], p = .034). Results are reported with 95% confidence intervals in accordance with APA guidelines.

---

# References

```
Moro, S., Rita, P., & Cortez, P. (2014). Bank marketing [Dataset]. 
  UCI Machine Learning Repository. https://doi.org/10.24432/C5K306

Moro, S., Cortez, P., & Rita, P. (2014). A data-driven approach to predict 
  the success of bank telemarketing. Decision Support Systems, 62, 22–31. 
  https://doi.org/10.1016/j.dss.2014.03.001

Wickham, H., & Grolemund, G. (2017). R for data science. O'Reilly Media.
  https://r4ds.had.co.nz

Tufte, E. R. (2001). The visual display of quantitative information (2nd ed.). 
  Graphics Press.

R Core Team. (2024). R: A language and environment for statistical computing. 
  R Foundation for Statistical Computing. https://www.R-project.org/

[Add your literature review sources here in APA format]
```

---

*End of Project — `r Sys.Date()`*
