---
title: "Introduction to Data Analytics 2"
author: "Ravina Gaikwad"
date: "`r Sys.Date()`"
output: 
  word_document:
    reference_docx: default
---

```{r machine-info, echo=FALSE, results='asis'}
cat("**Machine:**", paste0(Sys.info()['nodename'], " - ", Sys.info()['sysname'], " ", Sys.info()['release']), "\n\n")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Set machine info for document
machine_info <- paste0(Sys.info()['nodename'], ' - ', Sys.info()['sysname'], ' ', Sys.info()['release'])
```

# Part 1: Experiment Design

*Title*: Consumer Pseudo-Showrooming and Omni-Channel Product Placement Strategies

*Abstract*: Recent advances in information technologies (IT) have powered the merger of online and offline retail channels into one single platform. 
Modern consumers frequently switch between online and offline channels when they navigate through various stages of the decision journey, motivating multichannel sellers to develop omni-channel strategies that optimize their overall profit. 
This study examines consumers' cross-channel search behavior of "pseudo-showrooming," or the consumer behavior of inspecting one product at a seller's physical store before buying a related but different product at the same seller’s online store, 
and investigates how such consumer behavior allows a multichannel seller to achieve better coordination between its online and offline arms through optimal product placement strategies.

Participants in the study were grouped into the following categories: 

  - Where_bought: Where they ended up purchasing an item: bought at the store, bought online. 
  - Who_bought: If they bought from the same or a different retailer. 

Each participant was then measured on:

  - Money: how much money they spent in dollars on the product.
  - Time: how much time (in minutes) they spent looking at the product online. 

1)	What would be one possible null hypothesis based on this study?

**RG Answer**
- H₀: There is no difference in money spent between purchasing at a store versus online, regardless of retailer (same vs. different).

The null hypothesis would be that there is no real difference in how much money people spend or how much time they take, whether they buy from the same store they looked at or from a different store. Basically, we're saying that where people buy (online vs physical store) and whether they switch retailers doesn't actually change their spending or time patterns

2)	What would be one possible alternative hypothesis based on this study?

**RG Answer**
- H₁ = at least one of the group averages is different from the others.

The alternative hypothesis would be that there IS a difference - that people who do pseudo-showrooming (looking at one store but buying from another) actually do spend different amounts of money or time compared to people who just buy from the same place they looked. So we're predicting that the shopping behavior actually matters and creates different outcomes.

3)	Who are they sampling in this study?

**RG Answer**

- The researchers are looking at regular shoppers who use both physical stores and online shopping. From the data, it looks like 200 poeple participated. These people might be who go to a store to check out a product in person, and then decide whether to buy it there or go online to buy it (maybe from the same company or maybe from a competitor).

4)	Who is the intended population in this study?

**RG Answer**
The population they're trying to learn about is basically all modern shoppers who use multiple channels to shop - people who don't just shop online OR in stores, but mix both. This includes pretty much anyone who might browse in a physical store but then order online, or who compares prices across different retailers. So it's trying to understand shopping habits for most people in today's retail environment.

5)	Give an example of type 1 error based on this study (do not just define, explain in context how it might have happened here).

**RG Answer**
A Type 1 error is when we think we found something real but it's actually just a coincidence. In this study, that would mean the researchers conclude that pseudo-showrooming makes people spend less money, so they tell stores to change their strategies. But actually, maybe the people in the "online-different" group just happened to be more budget-conscious shoppers by random chance when they were collecting data. So stores would change their whole business strategy based on a pattern that doesn't really exist, which could waste money and hurt sales. It's like thinking you found a pattern when really it was just random luck in who happened to be in your study.

6)	Give an example of type 2 error based on this study (do not just define, explain in context how it might have happened here).

**RG Answer:**
A Type 2 error is when something real is happening but we miss it in our analysis. For this study, maybe pseudo-showrooming actually DOES make people more efficient shoppers who spend less time and money, which would be really useful information for stores to know. But if the researchers didn't have enough people in their study, or if people's shopping habits varied a lot, the statistics might not show a significant result. So the researchers would say "we didn't find anything" and stores would miss out on important insights about customer behavior. They'd keep doing business as usual when they could have been optimizing their strategy.

# Part 2: Use the 04_data.csv to complete this portion.

```{r import-data}
data <- read.csv("/Users/rgaikw468/Documents/Personal/R-Projects/500 HU 2025 vRG/03-Lab/04_data.csv")
head(data)
str(data)
```

1)	For each IV list the levels (next to a, b):

**RG Answer:**

a. Where bought: **online, store**

b. Who bought: **different, same**
    
2)	What are the conditions in this experiment?

**RG Answer:**
There are 4 different conditions because we have 2 independent variables with 2 levels each (2 x 2 = 4 conditions):

1. **Online + Different** - People who looked at a product and then bought a different product online
2. **Online + Same** - People who bought the same product they looked at, but online
3. **Store + Different** - People who looked at a product and bought a different one in the physical store
4. **Store + Same** - People who bought the same product they looked at, in the physical store

3)	For each condition list the means, standard deviations, and standard error for the conditions for time and money spent. Please note that means you should have several sets of M, SD, and SE. Be sure you name the sets of means, sd, and se different things so you can use them later.

```{r descriptives}
# 4 groups/conditions
online_different <- data[data$where_bought == "online" & data$who_bought == "different", ]
online_same <- data[data$where_bought == "online" & data$who_bought == "same", ]
store_different <- data[data$where_bought == "store" & data$who_bought == "different", ]
store_same <- data[data$where_bought == "store" & data$who_bought == "same", ]

# Now calculate statistics for Online + Different
n_online_diff <- nrow(online_different)
mean_money_online_diff <- mean(online_different$money)
sd_money_online_diff <- sd(online_different$money)
se_money_online_diff <- sd_money_online_diff / sqrt(n_online_diff)

mean_time_online_diff <- mean(online_different$time)
sd_time_online_diff <- sd(online_different$time)
se_time_online_diff <- sd_time_online_diff / sqrt(n_online_diff)

# Online + Same
n_online_same <- nrow(online_same)
mean_money_online_same <- mean(online_same$money)
sd_money_online_same <- sd(online_same$money)
se_money_online_same <- sd_money_online_same / sqrt(n_online_same)

mean_time_online_same <- mean(online_same$time)
sd_time_online_same <- sd(online_same$time)
se_time_online_same <- sd_time_online_same / sqrt(n_online_same)

# Store + Different
n_store_diff <- nrow(store_different)
mean_money_store_diff <- mean(store_different$money)
sd_money_store_diff <- sd(store_different$money)
se_money_store_diff <- sd_money_store_diff / sqrt(n_store_diff)

mean_time_store_diff <- mean(store_different$time)
sd_time_store_diff <- sd(store_different$time)
se_time_store_diff <- sd_time_store_diff / sqrt(n_store_diff)

# Store + Same
n_store_same <- nrow(store_same)
mean_money_store_same <- mean(store_same$money)
sd_money_store_same <- sd(store_same$money)
se_money_store_same <- sd_money_store_same / sqrt(n_store_same)

mean_time_store_same <- mean(store_same$time)
sd_time_store_same <- sd(store_same$time)
se_time_store_same <- sd_time_store_same / sqrt(n_store_same)

cat("ONLINE + DIFFERENT (n =", n_online_diff, "):\n")
cat("  Money: M =", round(mean_money_online_diff, 2), ", SD =", round(sd_money_online_diff, 2), ", SE =", round(se_money_online_diff, 2), "\n")
cat("  Time:  M =", round(mean_time_online_diff, 2), ", SD =", round(sd_time_online_diff, 2), ", SE =", round(se_time_online_diff, 2), "\n\n")

cat("ONLINE + SAME (n =", n_online_same, "):\n")
cat("  Money: M =", round(mean_money_online_same, 2), ", SD =", round(sd_money_online_same, 2), ", SE =", round(se_money_online_same, 2), "\n")
cat("  Time:  M =", round(mean_time_online_same, 2), ", SD =", round(sd_time_online_same, 2), ", SE =", round(se_time_online_same, 2), "\n\n")

cat("STORE + DIFFERENT (n =", n_store_diff, "):\n")
cat("  Money: M =", round(mean_money_store_diff, 2), ", SD =", round(sd_money_store_diff, 2), ", SE =", round(se_money_store_diff, 2), "\n")
cat("  Time:  M =", round(mean_time_store_diff, 2), ", SD =", round(sd_time_store_diff, 2), ", SE =", round(se_time_store_diff, 2), "\n\n")

cat("STORE + SAME (n =", n_store_same, "):\n")
cat("  Money: M =", round(mean_money_store_same, 2), ", SD =", round(sd_money_store_same, 2), ", SE =", round(se_money_store_same, 2), "\n")
cat("  Time:  M =", round(mean_time_store_same, 2), ", SD =", round(sd_time_store_same, 2), ", SE =", round(se_time_store_same, 2), "\n")
```
**RG Answer:**
You'll find Mean(M), Standard Deviation(SD) and Standard Error(SE) for each conditions per time and money spent below:

ONLINE + DIFFERENT (n = 50 ):
  Money: M = 25.38 , SD = 2.81 , SE = 0.4 
  Time:  M = 10.05 , SD = 2.75 , SE = 0.39 

ONLINE + SAME (n = 50 ):
  Money: M = 25.13 , SD = 3.26 , SE = 0.46 
  Time:  M = 10.52 , SD = 3.03 , SE = 0.43 

STORE + DIFFERENT (n = 50 ):
  Money: M = 34.7 , SD = 3.05 , SE = 0.43 
  Time:  M = 15.33 , SD = 2.66 , SE = 0.38 

STORE + SAME (n = 50 ):
  Money: M = 34.88 , SD = 2.74 , SE = 0.39 
  Time:  M = 20.1 , SD = 2.91 , SE = 0.41 



4)	Which condition appears to have the best model fit using the mean as the model (i.e. smallest error) for time?

```{r best-fit}
# Compare the standard errors for time
cat("Standard Errors for TIME across conditions:\n")
cat("Online + Different: SE =", round(se_time_online_diff, 2), "\n")
cat("Online + Same:      SE =", round(se_time_online_same, 2), "\n")
cat("Store + Different:  SE =", round(se_time_store_diff, 2), "\n")
cat("Store + Same:       SE =", round(se_time_store_same, 2), "\n")
```
**RG Answer:**

The condition with the smallest standard error for time is **Store + Different** with SE = `r round(se_time_store_diff, 2)`. **Store + Different** seems to be the best fit.

5)	What are the df for each condition?

```{r df}
# Degrees of freedom is always n - 1 for each group
df_online_diff <- n_online_diff - 1
df_online_same <- n_online_same - 1
df_store_diff <- n_store_diff - 1
df_store_same <- n_store_same - 1

cat("Degrees of Freedom (df = n - 1):\n")
cat("Online + Different: df =", df_online_diff, "\n")
cat("Online + Same:      df =", df_online_same, "\n")
cat("Store + Different:  df =", df_store_diff, "\n")
cat("Store + Same:       df =", df_store_same, "\n")
```
**RG Answer:**
Since degrees of freedom is n-1 for each group, df is as follows:
Online + Different: df = 49 
Online + Same:      df = 49 
Store + Different:  df = 49 
Store + Same:       df = 49 

6)	What is the confidence interval (95%) for the means?

```{r conf-interval}
##money

# For Online + Different
ci_lower_money_online_diff <- mean_money_online_diff - qt(0.975, df_online_diff) * se_money_online_diff
ci_upper_money_online_diff <- mean_money_online_diff + qt(0.975, df_online_diff) * se_money_online_diff

# For Online + Same
ci_lower_money_online_same <- mean_money_online_same - qt(0.975, df_online_same) * se_money_online_same
ci_upper_money_online_same <- mean_money_online_same + qt(0.975, df_online_same) * se_money_online_same

# For Store + Different
ci_lower_money_store_diff <- mean_money_store_diff - qt(0.975, df_store_diff) * se_money_store_diff
ci_upper_money_store_diff <- mean_money_store_diff + qt(0.975, df_store_diff) * se_money_store_diff

# For Store + Same
ci_lower_money_store_same <- mean_money_store_same - qt(0.975, df_store_same) * se_money_store_same
ci_upper_money_store_same <- mean_money_store_same + qt(0.975, df_store_same) * se_money_store_same

cat("95% Confidence Intervals for MONEY:\n")
cat("Online + Different: [", round(ci_lower_money_online_diff, 2), ",", round(ci_upper_money_online_diff, 2), "]\n")
cat("Online + Same:      [", round(ci_lower_money_online_same, 2), ",", round(ci_upper_money_online_same, 2), "]\n")
cat("Store + Different:  [", round(ci_lower_money_store_diff, 2), ",", round(ci_upper_money_store_diff, 2), "]\n")
cat("Store + Same:       [", round(ci_lower_money_store_same, 2), ",", round(ci_upper_money_store_same, 2), "]\n\n")

##time

# For Online + Different
ci_lower_time_online_diff <- mean_time_online_diff - qt(0.975, df_online_diff) * se_time_online_diff
ci_upper_time_online_diff <- mean_time_online_diff + qt(0.975, df_online_diff) * se_time_online_diff

# For Online + Same
ci_lower_time_online_same <- mean_time_online_same - qt(0.975, df_online_same) * se_time_online_same
ci_upper_time_online_same <- mean_time_online_same + qt(0.975, df_online_same) * se_time_online_same

# For Store + Different
ci_lower_time_store_diff <- mean_time_store_diff - qt(0.975, df_store_diff) * se_time_store_diff
ci_upper_time_store_diff <- mean_time_store_diff + qt(0.975, df_store_diff) * se_time_store_diff

# For Store + Same
ci_lower_time_store_same <- mean_time_store_same - qt(0.975, df_store_same) * se_time_store_same
ci_upper_time_store_same <- mean_time_store_same + qt(0.975, df_store_same) * se_time_store_same

cat("95% Confidence Intervals for TIME:\n")
cat("Online + Different: [", round(ci_lower_time_online_diff, 2), ",", round(ci_upper_time_online_diff, 2), "]\n")
cat("Online + Same:      [", round(ci_lower_time_online_same, 2), ",", round(ci_upper_time_online_same, 2), "]\n")
cat("Store + Different:  [", round(ci_lower_time_store_diff, 2), ",", round(ci_upper_time_store_diff, 2), "]\n")
cat("Store + Same:       [", round(ci_lower_time_store_same, 2), ",", round(ci_upper_time_store_same, 2), "]\n")

```
**RG Answer:**

95% Confidence Intervals for Money are as follows:
Online + Different: [ 24.58 , 26.18 ]
Online + Same:      [ 24.21 , 26.06 ]
Store + Different:  [ 33.84 , 35.57 ]
Store + Same:       [ 34.1 , 35.66 ]


95% Confidence Intervals for Time are as follows:
Online + Different: [ 9.27 , 10.83 ]
Online + Same:      [ 9.66 , 11.38 ]
Store + Different:  [ 14.58 , 16.09 ]
Store + Same:       [ 19.27 , 20.93 ]


7)	Use the MOTE library to calculate the effect size for the difference between money spent for the following comparisons (that means you'll have to do this twice):

```{r MOTE}
#install.packages("MOTE")
library(MOTE)
##Store versus online when bought at the same retailer

effect_same_retailer <- d.ind.t(
  m1 = mean_money_store_same,
  m2 = mean_money_online_same,
  sd1 = sd_money_store_same,
  sd2 = sd_money_online_same,
  n1 = n_store_same,
  n2 = n_online_same,
  a = 0.05
)

cat("For Store vs Online (when buying from SAME retailer)\n")
cat("Cohen's d =", round(effect_same_retailer$d, 3), "\n\n")

##Store versus online when bought at a different retailer

effect_diff_retailer <- d.ind.t(
  m1 = mean_money_store_diff,
  m2 = mean_money_online_diff,
  sd1 = sd_money_store_diff,
  sd2 = sd_money_online_diff,
  n1 = n_store_diff,
  n2 = n_online_diff,
  a = 0.05
)

cat("For Store vs Online (when buying from DIFFERENT retailer)\n")
cat("Cohen's d =", round(effect_diff_retailer$d, 3), "\n")

```
**RG Answer:**
Using MOTE we find that for Store vs Online-SAME retailer, cohen's d = 3.24 & for Store vs Online-DIFFERENT retailer, cohen's d = 3.18.


8)	What can you determine about the effect size in the experiment - is it small, medium or large?

**RG Answer:**

Considering Cohen's d, the guidelines are:
- Small effect: d = 0.2
- Medium effect: d = 0.5
- Large effect: d = 0.8

For Same Retailer - Store vs Online: Cohen's d = `r round(effect_same_retailer$d, 2)`

This is way bigger than 0.8, so it's a **very large effect**. This means there's a huge difference in how much money people spend when they buy in-store versus online, even when they're buying from the same company.

For Different Retailer - Store vs Online: Cohen's d = `r round(effect_diff_retailer$d, 2)`

This is also way above 0.8, so again it's a **very large effect**. Whether people are switching retailers or not, buying in a physical store leads to spending way more money than buying online.

Both of these effect sizes are really big, which means the difference between store and online shopping is not just statistically significant - it's also meaningful.

9)	How many people did we need in the study for each comparison?

```{r pwr}
#install.packages("pwr")
library(pwr)
##Store versus online when bought at the same retailer

power_same <- pwr.t.test(
  d = effect_same_retailer$d,
  sig.level = 0.05,
  power = 0.80,
  type = "two.sample",
  alternative = "two.sided"
)

cat("For Store vs Online (Same Retailer)\n")
cat("Sample size needed per group:", ceiling(power_same$n), "\n")
cat("Total sample size needed:", ceiling(power_same$n) * 2, "\n\n")

##Store versus online when bought at a different retailer
#considering power of 0.80 and significance level of 0.05
power_diff <- pwr.t.test(
  d = effect_diff_retailer$d,
  sig.level = 0.05,
  power = 0.80,
  type = "two.sample",
  alternative = "two.sided"
)

cat("For Store vs Online (Different Retailer)\n")
cat("Sample size needed per group:", ceiling(power_diff$n), "\n")
cat("Total sample size needed:", ceiling(power_diff$n) * 2, "\n\n")

cat("Our actual sample sizes:\n")
cat("Online groups: n =", n_online_same, "\n")
cat("Store groups: n =", n_store_same, "\n")

```
**RG Answer:**

According to the power analysis, we only needed about **`r ceiling(power_same$n)` people per group** (so `r ceiling(power_same$n) * 2` people total) to detect these effects with 80% power. Since we actually have 50 people per group (200 total), looks like we have way more than enough participants.

This makes sense because our effect sizes were so large - when there's a really big difference between groups, you don't need as many people to be confident you'll find it. Study seems to be "overpowered" which means we could have used fewer people and still found these results.